---
title: "(P)artial (in)variance specification search"
author: "Mark Lai"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    df_print: paged
vignette: >
  %\VignetteIndexEntry{(P)artial (in)variance specification search}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package implements the sequential specification search method for identifying parameters violating measurement invariance across groups. The method is discussed in, for example, [Yoon and Millsap (2007)](https://doi.org/10.1080/10705510701301677), and is a popular method for identifying a partial invariance model.

## Workflow

A suggested workflow for using the main function `pinSearch()` is

1. Define the [`lavaan`](https://www.lavaan.ugent.be/) syntax for the configural invariance model.
2. Conduct a `pinSearch()` with specific test statistic and stopping criteria. The default is to use modification indices as in Yoon and Millsap (2007), but the score test is also supported. The function returns a final partial invariance model and a list of noninvariant parameters in the search order.
3. Obtain effect size statistics for the noninvariant items.

## Example

Here we use the classic data set by Holzinger & Swineford, with the goal of searching for a partial strict invariance model.

```{r}
library(pinsearch)
library(lavaan)
mod <-
    " visual  =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed   =~ x7 + x8 + x9 "
# Output the final partial invariance model, and the noninvariant items
ps1 <-
    pinSearch(mod,
        data = HolzingerSwineford1939,
        group = "school", type = "residuals",
        inv_test = "mod"  # use modification indices for search (default)
    )
```

Here are the noninvariant parameters:

```{r}
ps1$`Non-Invariant Items`
```

And the final partial invariance model

```{r}
ps1$`Partial Invariance Fit` |>
    summary()
```

## Effect Size

The package supports the $d_\text{MACS}$ effect size statistics by [Nye and Drasgow (2011)](https://doi.org/10.1037/a0022955) for noninvariant items, and the generalization to an $f_\text{bias}$ statistic for more than two groups.

These can be obtained by setting the `effect_size = TRUE` argument:

```{r}
pinSearch(mod,
    data = HolzingerSwineford1939,
    group = "school", type = "residuals",
    inv_test = "mod",
    effect_size = TRUE
)
```

Note that in the above output, the effect sizes are printed three times for `x3` and `x7`, which have non-invariant intercepts. This is because there are three latent variables in the model, and the function allows for the possibility of cross-loadings. When each item is only loaded on one latent variable, user can ignore the duplicated values (we will try to fix this in future releases).

Another possibility is to use the `pin_effsize()` function, which also works for partial invariance model not obtained from `pinSearch()`.

```{r}
# Fit the partial strict model by hand
pstrict_fit <- cfa(mod, data = HolzingerSwineford1939,
                   group = "school",
                   group.equal = c("loadings", "intercepts",
                                   "residuals"),
                   group.partial = c("x3~1", "x7~1"))
# Effect sizes
pin_effsize(pstrict_fit)
```

## Cross-Loadings

The function also supports models with cross-loadings. For example,

```{r}
mod2 <- "
   visual =~ x1 + x2 + x3 + x9
  textual =~ x4 + x5 + x6
  speed   =~ x7 + x8 + x9
"
ps2 <-
    pinSearch(mod2,
        data = HolzingerSwineford1939,
        group = "school", type = "residuals",
        inv_test = "mod",
        effect_size = TRUE
    )
ps2
```

It just happened in this examplle that the final results are not affected.

## Score Test

One can also use the score test instead of the modification indices, as [starting in version 0.6-17, `lavaan` recommends using that for equality constraints](https://lavaan.ugent.be/history/dot6.html#version-0.6-17). This can be achieved with `inv_test = "score"`.

```{r}
pinSearch(mod2,
    data = HolzingerSwineford1939,
    group = "school", type = "residuals",
    inv_test = "score",
    effect_size = TRUE
)
```

The test is slightly more sensitive and identifies an additional non-invariant uniqueness for item `x7`.

## Likelihood Ratio Test (LRT)

LRT is generally more accurate, but more computationally expensive.

```{r}
pinSearch(mod2,
    data = HolzingerSwineford1939,
    group = "school", type = "residuals",
    inv_test = "lrt",
    effect_size = TRUE
)
```

In this example, the results using LRT are the same as the score test.

## Controlling for False Discovery Rate

The function also supports adjusting for multiplicity by using a more stringent threshold for flagging a parameter as noninvariant, using the multistage false discovery rate procedure by [Benjamini & Gavrilov (2009)](https://projecteuclid.org/journals/annals-of-applied-statistics/volume-3/issue-1/A-simple-forward-selection-procedure-based-on-false-discovery-rate/10.1214/08-AOAS194.pdf) for forward selection. It is up for debate whether controlling for false discovery rate is as important in measurement/factorial invariance analyses as in other context, and it probably depends on whether one is interested in identifying non-invariant items or in inferences for the latent parameters. If this is of interest, one can use the `control_fdr = TRUE` option.

```{r}
pinSearch(mod2,
    data = HolzingerSwineford1939,
    group = "school", type = "residuals",
    inv_test = "score",
    effect_size = TRUE,
    control_fdr = TRUE
)
```

The uniqueness for `x7` is no longer flagged.
